Non-word repetition in Yélî Dnye

### 2021-01-05 manuscript.Rmd
shorten YD phonology subsection

<!-- a2m No, I didn't propose that we would use NWR because I wanted to contribute to our understanding of the NWR task in new settings, but rather because it's quite a standard task so it could be useful for us documenting phonological development in Rossel. To make this clearer, I moved it down and integrated it with the four contributions below -->

@@

Yélî Dnye is an isolate language (presumed Papuan) spoken by approximately 7,000 people residing on Rossel Island, an island found at the far end of the Louisiade Archipelago in Milne Bay Province, Papua New Guinea. The Yélî sound system, much like its baroque grammatical system [@levinsonYDgrammar], is unlike any other in the region.

With only four primary places of articulation (bilabial, alveolar, post-alveolar, and velar) and no voicing contrasts, the phonological inventory is remarkably packed with acoustically similar segments. The core oral stop set includes both singleton (/p/, /t/, /ṭ/, and /k/) and doubly-articulated (/tp/, /ṭp/, /kp/) segments, with full nasal equivalents (/m/, /n/, /ṇ/, /ŋ/, /nm/, /ṇm/, /ŋm/), and with a substantial portion of them contrastively pre-nasalized or nasally released (/mp/, /nt/, /ṇṭ/, /ŋk/, /nmtp/, /ṇmṭp/, /ŋmkp/, /ṭṇ/, /kŋ/, /ṭpṇm/, /kpŋm/). A large number of this combinatorial set can further be contrastively labialized, palatalized on release, or both (e.g., /pʲ/, /pʷ/, /pʲʷ/; /tpʲ/; /ṇmḍbʲ/; see @levinsonYDgrammar for details).^[We use Levinson's [-@levinsonYDgrammar] under-dot notation (e.g., /ṭ/) to indicate the post-alveolar place of articulation; these stops are, articulatorily, somewhat variable in place, with at least some tokens produced fully sub-apically. In approximating cross-linguistic segment frequency below we use the corresponding retroflex for each stop segment (e.g., /ʈ/, /ʈp/, /ɳ/).] The consonantal inventory also includes a number of non-nasal continuants (/w/, /j/, /ɣ/, /l/, /βʲ/, /lʲ/, /lβʲ/). Vowels in Yélî Dnye may be oral or nasal, short or long. The 10 oral vowel qualities, which span four levels of vowel height, (/i/, /ɯ/, /u/, /e/, /o/, /ə/, /ɛ/, /ɔ/, /æ/, /ɑ/) can be produced as short and long vowels, with 7 of these able to appear as short and long nasal vowels as well /ĩ/, /ũ/, /ə̃/, /ɛ̃/, /ɔ̃/, /æ̃/, /ɑ̃/).

In total, Yélî Dnye uses 90 distinctive segments (not including an additional 3 rarely used consonants), far outstripping the phonemic inventory size of other documented Papuan languages [@foley1986papuan; @levinsonYDgrammar] and includes at least two contrasts that had not previously been documented elsewhere [labial-coronal double-articulations with dental vs. post-alveolar coronal placement for both oral and nasal stops, @ladefoged1996sounds; @maddiesonIPphoneticsYD].

Most words in Yélî Dnye are bisyllabic (~50%), with monosyllabic words (~40%) appearing most commonly after that, and with tri-and-above syllabic words appearing least frequently [~10%; based on >5800 lexemes in the most recent Yélî Dnye dictionary at the time of writing; @levinsonYDgrammar]. The vast majority of syllables use a CV format. A small portion of the lexicon features words with a final closed syllable, limited to codas of -/m/, -/p/, or -/j/ (e.g., "ndap" (Spondylus shell) /ṇṭæp/). However, in spontaneous speech an epenthetic /ɯ/ is often inserted after word-final coda, yielding a CV.CV structure instead (e.g., "ndap" pronounced "ndapî" /'ṇṭæ.pɯ/). This process is used frequently with English loan words that have a coda (e.g., "ponî" (phone) /'pɔ.nɯ/). An even smaller portion of the lexicon features words starting with a vowel (e.g., "ala" (here) /æ.'læ/), but these are limited to /æ/-. Finally, the lexicon features a handful of single-vowel grammatical morphemes (see @levinsonYDgrammar for details).

>>>>

@@

Yélî Dnye is an isolate language (presumed Papuan) spoken by approximately 7,000 people residing on Rossel Island, an island found at the far end of the Louisiade Archipelago in Milne Bay Province, Papua New Guinea. The Yélî sound system, much like its baroque grammatical system [@levinsonYDgrammar], is unlike any other in the region. In total, Yélî Dnye uses 90 distinctive segments (not including an additional 3 rarely used consonants), far outstripping the phonemic inventory size of other documented Papuan languages [@foley1986papuan; @levinsonYDgrammar; @maddiesonIPphoneticsYD]. <!--
Removed for brevity:
and includes at least two contrasts that had not previously been documented elsewhere [labial-coronal double-articulations with dental vs. post-alveolar coronal placement for both oral and nasal stops, @ladefoged1996sounds; @maddiesonIPphoneticsYD].
-->

With only four primary places of articulation (bilabial, alveolar, post-alveolar, and velar) and no voicing contrasts, the phonological inventory is remarkably packed with acoustically similar segments. The core oral stop system includes both singleton (/p/, /t/, /ṭ/, and /k/) and doubly-articulated (/tp/, /ṭp/, /kp/) segments, with full nasal equivalents (/m/, /n/, /ṇ/, /ŋ/, /nm/, /ṇm/, /ŋm/), and with a substantial portion of them contrastively pre-nasalized or nasally released (/mp/, /nt/, /ṇṭ/, /ŋk/, /nmtp/, /ṇmṭp/, /ŋmkp/, /ṭṇ/, /kŋ/, /ṭpṇm/, /kpŋm/). A large number of this combinatorial set can further be contrastively labialized, palatalized on release, or both (e.g., /pʲ/, /pʷ/, /pʲʷ/; /tpʲ/; /ṇmḍbʲ/; see @levinsonYDgrammar for details).^[We use Levinson's [-@levinsonYDgrammar] under-dot notation (e.g., /ṭ/) to denote the post-alveolar place of articulation; these stops are, articulatorily, somewhat variable in place, with at least some tokens produced fully sub-apically. In approximating cross-linguistic segment frequency below we use the corresponding retroflex for each stop segment (e.g., /ʈ/, /ʈp/, /ɳ/).] The consonantal inventory also includes a number of non-nasal continuants (/w/, /j/, /ɣ/, /l/, /βʲ/, /lʲ/, /lβʲ/). Vowels in Yélî Dnye may be oral or nasal, short or long. The 10 oral vowel qualities, which span four levels of vowel height, (/i/, /ɯ/, /u/, /e/, /o/, /ə/, /ɛ/, /ɔ/, /æ/, /ɑ/) can be produced as short and long vowels, with 7 of these able to appear as short and long nasal vowels as well /ĩ/, /ũ/, /ə̃/, /ɛ̃/, /ɔ̃/, /æ̃/, /ɑ̃/).

Most Yélî Dnye words are bisyllabic (~50%), with monosyllabic words (~40%) appearing most commonly after that, and with tri-and-above syllabic words appearing least frequently [~10%; based on >5800 lexemes in the most recent dictionary at the time of writing; @levinsonYDgrammar]. The vast majority of syllables use a CV format. A small portion of the lexicon features words with a final CVC syllable, but these are limited to codas of -/m/, -/p/, or -/j/ (e.g., "ndap" (Spondylus shell) /ṇṭæp/) and are often resyllabified with an epenthetic /ɯ/ in spontaneous speech (e.g., "ndapî" /'ṇṭæ.pɯ/). <!--Removed for brevity: This process is used frequently with English loan words that have a coda (e.g., "ponî" (phone) /'pɔ.nɯ/).--> There are also a handful of words starting with /æ/ (e.g., "ala" (here) /æ.'læ/) and a small collection of single-vowel grammatical morphemes (see @levinsonYDgrammar for details).
